Semi-Permeable Vesicles: Mathematical Formulation, Numerics, and Physics

### 2020-12-08 vesicle_code/docs/permeableVes/main.tex
updated shear language

We consider a semi-permeable vesicle in the shear flow
$\uu_{\infty}(\xx) = \dot{\gamma} (y,0)$ with different permeability
rates $\beta$ and shear rates $\dot{\gamma}$. Depending on its reduced
area and viscosity contrast, two-dimensional vesicles either tank tread
or tumble~\cite{fin-lam-sei-gom2008, kra-win-sei-lip1996}. In our
experiments, tumbling is impossible since we consider semi-permeable
vesicles with no viscosity contrast. Similar to the dynamics of an
impermeable vesicle, a semi-permeable vesicle tilts to an inclination
angle and undergoes tank treading dynamics.

In the quiescent flow, the steady state vesicle shape is independent of
the initial vesicle shape, and we investigate if this is true when a
shear flow is imposed. In Figure~\ref{fig:shearComposite}(a), using four
initial reduced areas and two flow rates, we plot the reduced area of a
semi-permeable vesicle with $\beta = 10^{-3}$. As we saw for the
quiescent example, the final shape does not depend on the initial
reduced area. However, it does depend on the flow rate. We also observe
a very small change in the reduced area at early times with the vesicle
dynamics at early times being similar to those of an impermeable
vesicle.  However, once the vesicle is tank-treading, the normal
component of the force-balance velocity is small and the permeability
velocity dominates and leads to a changing reduced area.

We also form the steady state shape for a variety of semi-permeability
rates and flow rates. Since the final shape is independent of the
initial reduce area, we choose an vesicle with $\nu = 0.65$ for all the
simulations. The steady state vesicle shape, along with its reduced
area, are in Figure~\ref{fig:shearComposite}(b). We see that the flow
rate significantly affects the final vesicle shape, but the
semi-permeability rate has no significant effect. Instead, as we saw in
the quiescent analysis, $\beta$ sets a time scale since the time
required for a vesicle to reach its steady state configuration is
proportional to $\beta^{-1}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Poiseuille Flow}
\begin{figure}[htp]
%  \ifTikz
  \input{figures/parabolicComposite.tikz}
%  \fi
  \caption{\label{fig:parabolicComposite} (a) The steady state shape of a
  semi-permeable vesicle submerged in a Poiseuille flow with varying
  initial reduced areas and flow rates. (b) The steady-state inclination
  angle for four different flow velocities. The steady-state shapes are
  superimposed. (c) The reduced area of the slipper formed with the
  flow rate of 800 $\mu m/s$. (d) The vesicle shape shown with the
  background imposed flow. The channel width is 12.5 times larger than
  the vesicle radius. The corresponding reduced area for each shape is
  indicated by the marks in plot (c).}
\end{figure}

>>>>

We consider a semi-permeable vesicle in the shear flow
$\uu_{\infty}(\xx) = \dot{\gamma} (y,0)$ with different permeability
rates $\beta$ and shear rates $\dot{\gamma}$. A vesicle's reduced area
and viscosity contrast determines whether it tank treads or
tumbles~\cite{fin-lam-sei-gom2008, kra-win-sei-lip1996}. Similar to the
dynamics of an impermeable vesicle, a semi-permeable vesicle tilts to an
inclination angle and undergoes tank treading dynamics. Since we
consider semi-permeable vesicles, there is no viscosity contrast and
tumbling is not possible.

In Figure~\ref{fig:shearComposite}(a), using four initial reduced areas
and two flow rates, we plot the reduced area of a semi-permeable vesicle
with $\beta = 10^{-3}$. As observed in a quiescent flow, the final
vesicle shape is independent of the initial reduced area. However, it
does depend on the flow rate. We also observe small changes in the
reduced area at early times with the vesicle dynamics at early times
that are similar to those of an impermeable vesicle. However, once the
vesicle is tank-treading, the normal component of the force-balance
velocity is small and the permeability velocity dominates and leads to a
changing reduced area.

We form the steady state shape for a variety of semi-permeability rates
and flow rates. Since the final shape is independent of the initial
reduce area, for all simulations we initialize the vesicle with reduced
area $\nu = 0.65$. The steady state vesicle shape and its reduced
area are in Figure~\ref{fig:shearComposite}(b). We see that the flow
rate significantly affects the final vesicle shape, but the
semi-permeability rate has a minor effect. Instead, as we saw in the
quiescent analysis, $\beta$ sets a time scale since the time required
for a vesicle to reach its steady state configuration is proportional to
$\beta^{-1}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Poiseuille Flow}
\begin{figure}[htp]
  \ifTikz
  \input{figures/parabolicComposite.tikz}
  \fi
  \caption{\label{fig:parabolicComposite} (a) The steady state shape of a
  semi-permeable vesicle submerged in a Poiseuille flow with varying
  initial reduced areas and flow rates. (b) The steady-state inclination
  angle for four different flow velocities. The steady-state shapes are
  superimposed. (c) The reduced area of the slipper formed with the
  flow rate of 800 $\mu m/s$. (d) The vesicle shape shown with the
  background imposed flow. The channel width is 12.5 times larger than
  the vesicle radius. The corresponding reduced area for each shape is
  indicated by the marks in plot (c).}
\end{figure}
